Dike experiment

### 2022-12-14 manuscript.Rmd
Start manuscript

---
title: "Dike experiment"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output:
  bookdown::word_document2:
    reference_docx: 2111_Wissenschaftliche_Arbeit.docx # template containing the formatting you want to have in your knitted word document
    # number_sections: false # only works for html or pdf...
    # using bookdown cross-referencing of sections possible, for flextable with pictures: officedown::rdocx_document
fontfamily: Arial
fontsize: 11pt
geometry: margin = 1in
csl: journal-of-vegetation-science.csl
---

Chair of Restoration Ecology, TUM School of Life Sciences, Technichal University of Munich, Germany

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de){.email}

@@

@@

```{r set.up, include = F}
# Create objects for easy access to the folders related to the project
figures_dir <- file.path("outputs", "figures")

# plot figures high resolution
knitr::opts_chunk$set(dpi = 600)

library(knitr) # importing figures

@@

# Abstract {.unnumbered}

@@

**keywords:** keyword1, keyword2, keyword3, keyword4, keyword5

# Introduction {.unnumbered}

Climate change is globally one of the main drivers restructuring species communities [@halsch2021; @parmesan2006; @hill2021]. [...] to answer the following questions:

-   Question 1?
-   Question 2?

# Material & Methods {.unnumbered}

## Study area {.unnumbered}

@@

## Data Collection {.unnumbered}

# Results {.unnumbered}

# Discussion {.unnumbered}

@@

@@

**Figure 1:**

Description Figure 1.

**Figure 2:** Description Figure 2.

**Figure 3:** Description Figure 3.

@@

**Figure 1:**

```{r chunk-figure1, fig.cap = "", echo=FALSE}
include_graphics(here::here(figures_dir, "figure_1_map_tmap_300dpi_8x11cm.tiff"))
```

@@

**Figure 2:**

```{r chunk-figure2, fig.cap = "", echo = FALSE}
include_graphics(here::here(figures_dir, "figure_2_800dpi_16.5x5cm.tiff"))
```

@@

**Figure S1:**

```{r chunk-supp-1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Legend Figure S1", dpi=600}
#include_graphics(here::here("text", "supplementary", "figureS1.png"))
```

@@

**Figure S2:**

```{r chunk-supp-2, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Legend Figure S2", dpi=600}
#include_graphics(here::here("text", "supplementary", "figureS2.png"))
```

@@

## Session Info

*R version, the OS and attached or loaded packages:*

>>>>

---
title: "Dike experiment"
author: Markus Bauer*, Jakob Huber, Johannes Kollmann
output: bookdown::html_document2
fontfamily: Times New Roman
fontsize: 11pt
geometry: margin = 1in
csl: template-journal-of-vegetation-science.csl
bibliography: bibliography.bib
github-repo: markus1bauer/2023_danube_dike_experiment
---

Chair of Restoration Ecology, TUM School of Life Sciences, Technical University of Munich, Germany

\* Corresponding author: [markus1.bauer\@tum.de](mailto:markus1.bauer@tum.de)

@@

**Max 8000 words**

@@

```{r set-up, include = F}
library(knitr)
knitr::opts_chunk$set(
  echo = FALSE,
  dpi = 100
  )

@@

# Abstract (max 300 words) {.unnumbered}

@@

# Keywords {.unnumbered}

dike grassland

embankment

species composition

environmental filter

# Introduction

"Ecological theory can point restoration toward important processes that need manipulation [...]. However, for this information to be relevant, restoration ecology needs to employ evidence-based assessments [...]" (Suding 2011) --\> No evidence for theory with our project

-   Endangered grasslands

-   Predictability of restorations --\> experiments, evidence-based restoration (cooke 2018)

-   Brudvig 2017 among restoration outcomes and in comparison to reference

-   Substrate modification to influence habitat filtering

-   Sowing a typical approach + priority effects + species pool

-   Introduce FCS, Recovery time, Persistence

-   Dike grasslands as case study

-   The aim of the study is

We asked following question to evaluate the restoration success after four years:

1.  How close is the vegetation to the reference state?

2.  How strong differs the Favourable Conservation Status (FCS) among the seed mixture-substrate combinations?

3.  How strong differs the persistence of the sown species among the seed mixture-substrate combinations?

# Material & Methods

## Experimental design {.unnumbered}

The seed mixture-substrate combination were tested with an experiment on a dike of the River Danube in SE Germany, which was established in March 2018 (Figure \@ref(fig:figure-1)), 314 m above sea level (asl); WGS84 (lat/lon), 48.83895, 12.88412). The climate is temperate-suboceanic with a mean annual temperature of 8.4 °C and precipitation of 984 mm (**Deutscher Wetterdienst [DWD], 2021**). During the study, three exceptionally dry years (2018--2020) occurred **(Climate Data Center of the German Meteorological Service [CDC], 2022a, 2022b**), and three minor floods which did not reach the plots (**Bayerisches Landesamt für Umwelt [LfU], 2021a, 2021b)** (Table \@ref(tab:table-s1)). Regional sand (0--4 mm) was used to lean the substrate and the soil was taken from the dike <mark>Is that right?</mark> and a with an excavator the substrates were mixed <mark>Is that right?</mark> and the plots prepared. The target vegetation were lowland hay meadows (Arrhenatherion elatioris, CM01A) and calcareous grassland (Cirsio-Brachypodion pinnati, DA01B code of the EuroVegChecklist, **Mucina et al., 2016**). The species pool of hay meadows and calcareous grasslands consisted of 55 and 58 species <mark>check numbers</mark>, respectively. The seeds were received from a commercial producer of autochtonous seeds. From the species pools, 20 species were randomly selected for each plot. The seed mixtures always contained seven grasses (60 wt% of total seed mixture), three legumes (5%) and ten herbs (35%) (Table \@ref(tab:table-2))<mark>check</mark>. The community-weighted means (CWM) of functional traits differed between the seed mixture types (Table \@ref(tab:table-3)) <mark>more details and check</mark>. The management started with a cut without hay removal and a mowing height of 20 cm in August 2018, followed by 'normal' deep cuts with hay removal in July 2019 and 2020. In October 2018, Bromus hordeaceus was seeded as a nursery grass to provide safe sites under drought conditions.

We used 288 plots of the size 2.0 m × 2.0 m which were distributed over the north and south exposition of the dike and arranged in six blocks (= replicates). The experiment used a split plot design combined with randomized complete block design (Figure \@ref(fig:figure-1)). The split plot was cuased by testing all 24 treatments on both sides of the dike. We tested three substrates with 0%, 25% and 50% sand admixture and two soil depths (15 vs. 30 cm). Below the substrate, a drainage layer of 5 cm consisting of gravel (0--16 mm) was installed. The sand admixture changed the soil texture, increased the C/N ratio, but reduced the ratio of Calciumcarbonat, and hardly changed the pH (Table \@ref(tab:table-1)). The two seed mixture types were tested with two seeding densities (4 vs. 8 g m-2). The vegetation was surveyed in June or July 2018--2021 (**Braun-Blanquet, 1928/1964**). Establishment rates of species

We performed all analyses in R (Version 4.2.2; @base), with the functions 'blme' (based on 'lme4') for BLMM [@lme4] **(Bates, Mächler, Bolker, & Walker, 2015; Chung, Rabe-Hesketh, Dorie, Gelman, & Liu, 2013)**; and 'DHARMa' for model evaluation [@DHARMa].

<mark>We calculated Bayesian linear mixed-effects models (BLMM) with the random effect 'plot' and used the restricted maximum-likelihood estimation (REML), the optimiser Nelder-Mead and, for the random effect, the Wishart prior. To identify the final model, we first reviewed the residual diagnostics of the candidate models and subsequently compared the remaining models using the Akaike information criterion adjusted for a small sample size (AICc) and chose the most parsimonious model (Appendix A4). Finally, we calculated the marginal and conditional coefficients of determination (R²m, R²c) and the 95% confidence intervals of the response variables.</mark>

# Results

Figure @ref(fig:figure-1)

@teixeira2022

# Discussion

@@

@@

We would like to thank our project partners Dr. Markus Fischer, Frank Schuster, and Christoph Schwahn (WIGES GmbH) for numerous discussions on restoration and management of dike grasslands. Field work was supported by Clemens Berger and Uwe Kleber-Lerchbaumer (Wasserwirtschaftsamt Deggendorf). <mark>Mention Naturschutzbehörde and funding from WIGES.</mark> We thank Holger Paetsch, Simon Reith, Anna Ritter, Jakob Strak, Leonardo H. Teixeira, and Linda Weggler for assisting with the field surveys or soil analyses in 2018--2020. The German Federal Environmental Foundation (DBU) supported MB with a doctoral scholarship.

# Author contribution {.unnumbered}

JH and JK designed the experiment. JH did the surveys in the years 2018--2020 and MB in 2019 and 2021. MB did the analyses and wrote the manuscript. JK and JH critically revised the manuscript.

# Open research {.unnumbered}

Data and code is available on Zenodo: <mark><https://doi.org/10.XXXX></mark>

Model evaluation is stored on GitHub: [github.com/markus1bauer/2023_danube_dike_experiment](https://github.com/markus1bauer/2023_danube_dike_experiment/tree/main/markdown)

# Funding {.unnumbered}

MB is funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020.

@@

# Tables {.unnumbered}

## Table 1 {.unnumbered}

```{r table-1, fig.cap = ""}
#include_graphics(here::here("outputs", "tables", "figureS1.png"))
```

@@

## Table 2 {.unnumbered}

```{r table-2, fig.cap = ""}
#include_graphics(here::here("outputs", "tables", "figureS1.png"))
```

@@

## Figure 1 {.unnumbered}

```{r figure-1, fig.cap = "figure-1"}
knitr::include_graphics(here::here("outputs", "figures", "figure_1_map_tmap_300dpi_8x11cm.tiff"))
```

## Figure 2 {.unnumbered}

```{r figure-2, fig.cap = ""}
knitr::include_graphics(here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff"))
```

## Figure 3 {.unnumbered}

```{r figure-3, fig.cap = ""}
knitr::include_graphics(here::here("outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x14cm.tiff"))
```

## Figure 4 {.unnumbered}

```{r figure-4, fig.cap = ""}
knitr::include_graphics(here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x14cm.tiff"))
```

## Figure 5 {.unnumbered}

```{r figure-5, fig.cap = ""}
knitr::include_graphics(here::here("outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x14cm.tiff"))
```

Fig. 5:

## Figure 6 {.unnumbered}

```{r figure-6, fig.cap = ""}
knitr::include_graphics(here::here("outputs", "figures", "figure_6_nmds_800dpi_16.5x16cm.tiff"))
```

Fig. 6:

@@

## Table S1 {.unnumbered}

Three dry years (2018--2020) and three minor floods (2018, 2019, 2021) occurred during the study period. Annual temperature, precipitation and discharge of River Danube near the study sites (2017--2021), based on weather station Metten (mean, 1981--2010; ID, 3271; WGS84 (lat/lon), 48.85476 and 12.918911; CDC, 2022a, 2022b) and stream gauge Pfelling (ID, 10078000; WGS84 (lat/lon), 48.87975 and 12.74716; LfU, 2021a). HQ2 = Highest discharge with a probability of occurrence every second year. HSW = Highest water level for shipping.

```{r table-s1}
#knitr::include_graphics(here::here("outputs", "tables", "figureS1.png"))
```

@@

## Table S2 {.unnumbered}

```{r table-s2}
#knitr::include_graphics(here::here("outputs", "tables", "figureS2.png"))
```

@@

## Table S3 {.unnumbered}

```{r table-s3}
#knitr::include_graphics(here::here("outputs", "tables", "table_a3_seedmixes.csv"))
```

## Session Info {.unnumbered}
